---
title: "Circus Maximus - Baseline Playtest Analysis"
author: "Playtesting Report"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
    highlight: tango
  pdf_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6
)

# Load libraries
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(kableExtra)

# Source data loading functions
source("load-data.R")
```

# Executive Summary

This report analyzes the baseline playtest results to establish current game balance and identify areas for improvement.

```{r load-data}
# Load data (update path as needed)
df <- load_playtest_csv("../../data/exports/baseline.csv")
# Or from JSON:
# df <- load_playtest_json("../../data/raw/baseline/baseline-2024-01-15-100games-2players.json")

cat("Total games:", nrow(df), "\n")
cat("Completed games:", sum(df$completed), "\n")
cat("Completion rate:", round(mean(df$completed) * 100, 1), "%\n")
```

# Game Length Analysis

## Distribution of Game Length

```{r game-length-stats}
round_stats <- df %>%
  filter(completed == TRUE) %>%
  summarise(
    mean = mean(rounds),
    median = median(rounds),
    sd = sd(rounds),
    min = min(rounds),
    max = max(rounds),
    q25 = quantile(rounds, 0.25),
    q75 = quantile(rounds, 0.75)
  )

kable(round_stats, digits = 2, caption = "Game Length Statistics (Rounds)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r game-length-viz}
df %>%
  filter(completed == TRUE) %>%
  ggplot(aes(x = rounds)) +
  geom_histogram(binwidth = 1, fill = "steelblue", alpha = 0.7, color = "white") +
  geom_vline(aes(xintercept = mean(rounds)), color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = median(rounds)), color = "darkgreen", linetype = "dashed", size = 1) +
  labs(
    title = "Game Length Distribution",
    subtitle = paste("Mean =", round(mean(df$rounds[df$completed]), 1), "rounds | Median =", median(df$rounds[df$completed]), "rounds"),
    x = "Number of Rounds",
    y = "Frequency",
    caption = "Red line = Mean, Green line = Median"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"))
```

**Interpretation:**
- Target game length: 6-10 rounds
- Current mean: `r round(round_stats$mean, 1)` rounds
- **Verdict:** `r if(round_stats$mean < 6) {"Too short - consider increasing costs"} else if(round_stats$mean > 10) {"Too long - consider decreasing costs"} else {"Within acceptable range"}`

# Victory Track Balance

## Win Distribution by Track

```{r track-wins}
track_wins <- df %>%
  filter(completed == TRUE, !is.na(win_track)) %>%
  count(win_track, name = "wins") %>%
  mutate(
    win_rate = wins / sum(wins),
    expected_rate = 1/3,  # Expected 33.3% for balanced game
    difference = win_rate - expected_rate
  ) %>%
  arrange(desc(win_rate))

kable(track_wins, digits = 3, caption = "Victory Track Win Rates") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  add_header_above(c(" " = 1, "Observed" = 2, "Expected" = 1, "Difference" = 1))
```

```{r track-wins-viz}
track_wins %>%
  ggplot(aes(x = reorder(win_track, win_rate), y = win_rate)) +
  geom_col(fill = "steelblue", alpha = 0.7) +
  geom_hline(yintercept = 1/3, color = "red", linetype = "dashed", size = 1) +
  geom_text(aes(label = paste0(round(win_rate * 100, 1), "%")), 
            hjust = -0.1, size = 4) +
  coord_flip() +
  labs(
    title = "Victory Track Win Rates",
    subtitle = "Red line shows expected 33.3% for balanced game",
    x = "Victory Track",
    y = "Win Rate",
    caption = paste("Total games:", sum(track_wins$wins))
  ) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, max(0.5, max(track_wins$win_rate) * 1.1))) +
  theme_minimal()
```

**Interpretation:**
- Target: Each track should win ~33.3% of games (within ±10%)
- Largest imbalance: `r track_wins$win_track[1]` track wins `r round(track_wins$win_rate[1] * 100, 1)`% of games
- **Verdict:** `r if(abs(track_wins$difference[1]) > 0.1) {paste0("IMBALANCED - ", track_wins$win_track[1], " track is too strong/weak")} else {"Acceptably balanced"}`

## Statistical Test for Track Balance

```{r track-balance-test}
# Chi-square test for equal distribution
observed <- track_wins$wins
expected <- rep(sum(observed) / 3, 3)
chisq_result <- chisq.test(observed, p = rep(1/3, 3))

cat("Chi-square test for equal track distribution:\n")
cat("  Chi-square =", round(chisq_result$statistic, 3), "\n")
cat("  p-value =", format.pval(chisq_result$p.value, digits = 3), "\n")
cat("  Result:", if(chisq_result$p.value < 0.05) "IMBALANCED (p < 0.05)" else "No significant imbalance detected\n")
```

# Player Score Analysis

## Final Score Distribution

```{r player-scores}
df_long <- wide_to_long(df)

score_stats <- df_long %>%
  filter(completed == TRUE) %>%
  group_by(player_name) %>%
  summarise(
    mean_total = mean(total),
    mean_empire = mean(empire),
    mean_population = mean(population),
    mean_church = mean(church),
    win_rate = mean(won, na.rm = TRUE),
    n_games = n()
  ) %>%
  arrange(desc(mean_total))

kable(score_stats, digits = 2, caption = "Average Scores by Player") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r score-distribution}
df_long %>%
  filter(completed == TRUE) %>%
  ggplot(aes(x = total, fill = win_track_final)) +
  geom_histogram(binwidth = 2, alpha = 0.7, color = "white") +
  facet_wrap(~win_track_final, ncol = 1) +
  labs(
    title = "Final Score Distribution by Winning Track",
    x = "Total Track Score",
    y = "Frequency",
    fill = "Winning Track"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

# Confidence Intervals

```{r confidence-intervals}
# 95% confidence interval for mean game length
n_completed <- sum(df$completed)
mean_rounds <- mean(df$rounds[df$completed])
sd_rounds <- sd(df$rounds[df$completed])
se_rounds <- sd_rounds / sqrt(n_completed)
ci_rounds <- mean_rounds + c(-1, 1) * qt(0.975, n_completed - 1) * se_rounds

cat("Game Length - 95% Confidence Interval:\n")
cat(sprintf("  Mean: %.2f rounds\n", mean_rounds))
cat(sprintf("  95%% CI: [%.2f, %.2f] rounds\n", ci_rounds[1], ci_rounds[2]))
```

# Outlier Detection

```{r outliers}
# Identify outlier games (2+ standard deviations from mean)
mean_rounds <- mean(df$rounds[df$completed])
sd_rounds <- sd(df$rounds[df$completed])
threshold <- 2 * sd_rounds

outliers <- df %>%
  filter(completed == TRUE) %>%
  mutate(
    z_score = abs(rounds - mean_rounds) / sd_rounds,
    is_outlier = z_score > 2
  ) %>%
  filter(is_outlier) %>%
  select(game_id, rounds, winner, win_track, z_score)

if(nrow(outliers) > 0) {
  cat("Outlier games (>2 SD from mean):\n")
  kable(outliers, digits = 2) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
} else {
  cat("No outliers detected (>2 SD from mean)\n")
}
```

# Recommendations

## Balance Issues Identified

1. **Game Length:** `r if(round_stats$mean < 6) {"Games are too short. Consider increasing costs or victory threshold."} else if(round_stats$mean > 10) {"Games are too long. Consider decreasing costs or victory threshold."} else {"Game length is acceptable."}`

2. **Track Balance:** `r if(abs(track_wins$difference[1]) > 0.1) {paste0(track_wins$win_track[1], " track is winning ", round(track_wins$win_rate[1] * 100, 1), "% of games. Adjust track advancement rates or act card rewards.")} else {"Tracks are acceptably balanced."}`

3. **Completion Rate:** `r if(mean(df$completed) < 0.95) {paste0("Only ", round(mean(df$completed) * 100, 1), "% of games completed. Investigate infinite loop or max turn issues.")} else {"Most games completed successfully."}`

---

*Report generated on `r Sys.Date()`*
